\section{Introduction}
\label{sec:intro}

Agent-based models are used as policy instruments in epidemic
response, systemic risk, and other settings where heterogeneous
interaction is itself the object of study and a closed-form likelihood
is unavailable. Their calibration --- matching simulator output to
data without a likelihood --- has long been the bottleneck.
Differentiable agent-based modelling
\citep{querabofarull_diffabm_2023,querabofarull_diffabm_2025} has made
it tractable: the simulator is written as a chain of
automatic-differentiation primitives, so gradients of any discrepancy
loss are obtained in a single backward pass and the calibrator arrives
at every update with a buffer of per-seed gradients in hand. The
buffer carries information the calibrator has not yet been reading.

Yet first-order optimisers silently fail on these problems. With its
default hyperparameters, Adam diverges from $\theta^*$ on
Brock--Hommes calibration along a specific direction the loss does not
constrain, reaching the same MMD noise floor as a converged run while
wandering off in the unconstrained subspace. The available diagnostics
for such failure are either an offline finite-difference Hessian
\citep{naumann_woleske_2024}, too expensive to run during calibration,
or per-parameter Jacobian sensitivities
\citep{querabofarull_diffabm_2025}, which cannot detect parameter
\emph{combinations}. The culprit is non-identifiability of those
combinations under the MMD loss.

\citet{kunstner_opg_2019} pointed out that the outer-product-of-gradients
matrix differs from the Fisher under likelihood maximisation and can
mislead curvature-adaptive optimisers. We read the same matrix as a
residual generalised Gauss--Newton approximation of the MMD-loss
Hessian; the U-statistic structure of the MMD estimator licenses this
reading independently of any likelihood claim. The failure mode
Kunstner identified --- gradient-noise covariance misdirecting an
adaptive optimiser --- appears in non-likelihood agent-based
calibration as well, and it is detectable from the per-seed gradients
the calibrator already holds.

The contribution is a single live diagnostic, evidenced by three
independent checks. The eigendecomposition of the OPG matrix,
computed from the per-seed gradient buffer at no marginal cost, names
the parameter combinations the data cannot constrain before the next
update is taken (\S\ref{sec:method}, Figure~\ref{fig:method}). Three
discrepancies the calibrator never saw --- chosen to break
circularity with the MMD training loss --- agree with the diagnostic
across three structurally different models, with stiff/sloppy ratios
from $\mathbf{489\times}$ to $\mathbf{9.3 \times 10^5}$
(\S\ref{sec:results:falsification}, Figure~\ref{fig:falsification}).
The diagnostic predicts the direction of Adam's failure: across $25$
seeds spanning a financial heterogeneous-agent model and a mean-field
epidemic model, divergence runs along the flagged sloppy combination,
with cross-optimiser recovery-error ratio $\mathbf{9.4 \times 10^4}$
(\S\ref{sec:results:predicts-adam}, Figure~\ref{fig:predicts-adam}).
And the diagnostic survives the discrete-state regime: on a network
SIR model with Gumbel-Sigmoid surrogate gradients, the eigenstructure
remains qualitatively correct and the held-out ratios grow rather
than shrink (\S\ref{sec:results:network-sir},
Figure~\ref{fig:network-sir}).
