\section{Discussion}
\label{sec:discussion}

The outer-product-of-gradients matrix is two decades old as a
curvature surrogate
\citep{schraudolph_sgn_2002,martens_kfac_2015}; we use it for a
different purpose. For differentiable agent-based calibration, the
productive reading of $\widehat F$ is not as a preconditioner that
out-runs Adam but as a name for the parameter combinations the data
cannot constrain, available before the next update at the cost of one
eigendecomposition. The optimiser-replacement framing does not
survive the experiment. OPG-preconditioned Levenberg--Marquardt and
SGD reach indistinguishable loss floors on Brock--Hommes, and the
diagnostic predicts that neither will recover $\theta^*$ along the
sloppy directions; the same matrix tells us why.

The eigenvectors recover parameter combinations a domain expert would
derive from the model equations. On Brock--Hommes the stiff direction
is the symmetric-bias combination $(b_1 + b_2)/\sqrt{2}$ --- a
structure the per-parameter Jacobian sensitivities of
\citet{querabofarull_diffabm_2025} cannot surface. On SIR the sloppy
direction is the early-weak-versus-late-strong lockdown degeneracy
$f_{\mathrm{lock}} + 0.1\, t_{\mathrm{lock}}$, a public-health analogue
of the timing--intensity confounding any field epidemiologist will
recognise a fortnight into an outbreak. Structurally correct
combinations fall out of $\widehat F$ without being put in --- evidence
that the diagnostic recovers real identifiability structure, not
numerical artefact.

The network-SIR setting is the hardest of the three by construction.
Discrete-state transitions are reparameterised through a Gumbel-Sigmoid
surrogate \citep{maddison_concrete_2017,jang_gumbel_2017} whose
gradients are biased estimators of the true response, exactly the
regime in which $\widehat F$ might be expected to mislead. The
eigenvectors hold their qualitative shape regardless, and the non-MMD
falsification ratios grow rather than shrink under the bias.
Surrogate-gradient bias acts on the magnitudes of $\widehat F$, not on
the identifiability structure it carries.

Identifying the parameter combinations the data cannot constrain is a
safety property of a calibration pipeline: a diagnostic that flags an
overconfident calibration is more useful than one that silently
passes. The same matrix that delivers the optimisation step delivers
this flag, on every iterate, at the cost of a single eigendecomposition.
