\section{Limitations}
\label{sec:limitations}

The claims of this paper are scoped by four observations.

\paragraph{Horizon-dependence of magnitudes.}
The eigenvalue \emph{magnitudes} of $\widehat F$ depend on the gradient horizon $H$ used when backpropagating through the simulator. The qualitative stiff/sloppy hierarchy is stable across $H \in \{5, \ldots, 200\}$, but absolute eigenvalues shift up to $70\times$ between truncated and full-horizon gradients; reported spectrum spans should be matched on $H$.

\paragraph{Sample size of optimiser comparisons.}
The cross-optimiser results in Section~\ref{sec:results:predicts-adam} use $15$ Brock--Hommes seeds and $10$ SIR seeds. The ``$25/25$ Adam divergences'' claim has a 95\% lower confidence bound of approximately $88\%$, not asymptotic certainty; larger seed counts would tighten the bound.

\paragraph{Scale-confounding on SIR.}
The OPG spectrum span on the SIR models confounds eigenvalue with parameter scale (e.g.\ $I_0 \sim 10^{-3}$ versus $\beta \sim 1$). A Pearson-normalised correlation-matrix version of the spectrum gives a unit-free reading; we leave this to a future extension.

\paragraph{Optimiser claims are narrow.}
We claim that OPG-preconditioned Levenberg-Marquardt is robust without per-problem tuning on the two models tested, not that it beats best-tuned SGD in general; on Brock--Hommes the two are statistically tied at the best SGD learning rate.

Identifying the parameter combinations the data cannot constrain is a safety property: a diagnostic that flags an overconfident calibration is more useful than one that silently passes.
